---
title: "HW5_code"
format: 
  html:
    toc: true
    toc-location: left
    code-fold: true
    theme: yeti
execute: 
  message: false
  warning: false
editor_options: 
  chunk_output_type: console
---

## Homework 5 Code

```{r}
#install.packages("car")
```


```{r}
library(tidyverse)
library(here)
library(janitor)
library(ggeffects)
library(performance)
library(naniar) # or equivalent
library(flextable) # or equivalent
library(car)
library(broom)
```

```{r}
#install.packages("corrplot")
#install.packages("AICcmodavg")
#install.packages("GGally")
```

```{r}
library(corrplot)
library(AICcmodavg)
library(GGally)
```


```{r}
#reading in data table
plant<- read_csv(here("data", "knb-lter-hfr.109.18", "hf109-01-sarracenia.csv")) %>%
  #making the column names cleaner
  clean_names() %>% 
  select(totmass, species, feedlevel, sla, chlorophyll, amass, num_lvs, num_phylls)

head(plant)
```

```{r}
#missing data visualization
gg_miss_var(plant)

```

```{r}
plant_subset<- plant %>% 
  drop_na(sla, chlorophyll, amass, num_lvs, num_phylls)
view(plant_subset)
```

```{r}
#correlation plot
#To determine the relationships between numerical variables in our dataset, we calculated Pearsons r and visually represented correlation using a correlation plot.
# calculate Pearson's r for numerical values only
library(car)
plant_cor<-plant_subset %>% 
  select(feedlevel:num_phylls) %>% 
  cor(method="pearson")




```

```{r}
# creating a correlation plot
corrplot(plant_cor,
         # change the shape of what's in the cells
         method = "ellipse",
         addCoef.col = "black"
         )
```

```{r}
#creating a plot of each variable compared against the others
plant_subset %>% 
  select(species:num_phylls) %>% 
  ggpairs()
```

```{r}
#starting regression here
#To determine how species and physiological characteristics predict biomass, we fit multiple linear models.
null<- lm(totmass~1, data= plant_subset)
full <- lm(totmass ~ species + feedlevel + sla + chlorophyll + amass + num_lvs + num_phylls, data = plant_subset)
```

```{r}
# Diagnostics

#We visually assess normality and homoskedasticity of residuals using diagnostic plots for the full model
par(mfrow= c(2,2))
plot(full)
```

```{r}
#testing normality
check_normality(full)
check_heteroscedasticity(full)

```

```{r model-logs}
null_log <- lm(log(totmass) ~ 1, data = plant_subset)
full_log <- lm(log(totmass) ~ species + feedlevel + sla + chlorophyll + amass + num_lvs + num_phylls, data = plant_subset)

plot(full_log)
check_normality(full_log)
check_heteroscedasticity(full_log)
```

```{r}
#evaluate multicollinearity
car:: vif(full_log)
```
 
 we evaluated multicollinearity by calculating generalized variance inflation factor and determined that
 
 
 
 Trying some more models:
 
 what set of predictor variables best explains the response
 
```{r}
#second model
model2_log<- lm(log(totmass) ~ species, data= plant_subset)
```
 
```{r}
#check the plot
plot(model2_log)

check_normality(model2_log)
check_heteroscedasticity(model2_log)
```

compare models using Akaike's Information criterion (AIC) values:

```{r}
#AIC is useful for comparing 2 or more models. The AIC value comparison can tell us what model is the least complex and best predicts our response. Lowest AIC value means best model
AICc(full_log)
AICc(model2_log)
AICc(null_log)
```

```{r}
MuMIn::AICc(full_log, model2_log, null_log)
MuMIn:: model.sel(full_log, model2_log, null_log)
```



We found that the blank model, inlcuding the blank blank blank predictors best predicted blank (model summary).

```{r}
summary(full_log)

table <- tidy(full_log, conf.int = TRUE, exponentiate = TRUE) %>% 
  # change the p-value numbers if they're really small
  # change the estmaes, standard error, and t-tstatistics to round to ___ digits
  # using mutate
  # make it into a flextable
  flextable() %>% 
  # fit it to the viewer
  autofit()

table
```

```{r}
#using ggpredict to backtransform estimates 
model_pred <- ggpredict(full_log, terms = "species", back.transform = TRUE)

plot(ggpredict(full_log, terms = "species", back.transform = TRUE), add.data = TRUE)

plot(ggpredict(full_log, terms = "chlorophyll", back.transform = TRUE), add.data = TRUE)

plot(ggpredict(full_log, terms = "sla", back.transform = TRUE), add.data = TRUE)

model_pred

```
